Optimal (Multiway) Spatial Joins

### 2024-06-01 msj.tex
yufei

\def\xleft{\mathrm{left}}
\def\xright{\mathrm{right}}
\def\ybot{\mathrm{bot}}
\def\ytop{\mathrm{top}}
\def\maxtop{\mathrm{maxtop}}
\def\minleft{\mathrm{minleft}}
\def\gleft{\mathrm{left\text{-}guard}}
\def\gbot{\mathrm{bot\text{-}guard}}
\def\cross{\mathrm{cross}}
\def\dcross{\mathrm{d\text{-}cross}}
\def\contained{\mathrm{contained}}
\def\cat{\mathrm{cat1}}
\def\catt{\mathrm{cat2}}
\def\out{\mathrm{OUT}}
\def\dim{\mathrm{dim}}

@@

\section{The Comparison-Based Class for $k$-SJ} \label{app:comp-class}

Fix an integer $k \ge 3$, as well as $k$ integers $n_1, n_2, ..., n_k$ at least 1. Define $n = \sum_{i=1}^k n_i$. An {\em $(n_1, n_2, ..., n_k)$-input} to the $k$-SJ problem comprises $k$ sets of rectangles $R_1, R_2, ..., R_k$ in $\real^2$.

@@

A {\em comparison-based $k$-SJ} algorithm for solving $(n_1, n_2, ..., n_k)$-inputs is a decision tree $\T$ defined as follows. Each internal node $u$ of $\T$ has three child nodes, referred as the {\em left}, {\em middle}, and {\em right} child, respectively. Each node is associated with an {\em instruction} of the form $(i_1, \dim_1, \lambda_1; i_2, \dim_2, \lambda_2)$. Each leaf is associated with an {\em output}, which is defined to be a subset of $[n_1] \times [n_2] \times ... \times [n_k]$.

@@

Given an $(n_1, n_2, ..., n_k)$-input, we execute the algorithm from the root of $\T$.

>>>>

\def\xleft{\mathrm{left}}
\def\xright{\mathrm{right}}
\def\ybot{\mathrm{bot}}
\def\ytop{\mathrm{top}}
\def\maxtop{\mathrm{maxtop}}
\def\minleft{\mathrm{minleft}}
\def\gleft{\mathrm{left\text{-}guard}}
\def\gbot{\mathrm{bot\text{-}guard}}
\def\cross{\mathrm{cross}}
\def\dcross{\mathrm{d\text{-}cross}}
\def\contained{\mathrm{contained}}
\def\cat{\mathrm{cat1}}
\def\catt{\mathrm{cat2}}
\def\out{\mathrm{OUT}}
\def\side{\mathrm{side}}

@@

\section{Comparison-Based $k$-SJ Algorithms} \label{app:comp-class}

Fix an integer $k \ge 3$, as well as $k$ integers $n_1, n_2, ..., n_k$ at least 1. Define $n = \sum_{i=1}^k n_i$. An {\em $(n_1, n_2, ..., n_k)$-input} to the $k$-SJ problem comprises $k$ sets of rectangles $R_1, R_2, ..., R_k$ in $\real^2$ such that $|R_i| = n_i$ for each $i \in [k]$. We assume that each rectangle in $R_i$ ($i \in [k]$) is given a distinct ID from $[n_i]$.

@@

A {\em comparison-based $k$-SJ} algorithm for solving $(n_1, n_2, ..., n_k)$-inputs is a decision tree $\T$ defined as follows. Each internal node $u$ of $\T$ has three child nodes, referred as the {\em left}, {\em middle}, and {\em right} child, respectively. Each node is associated with an {\em instruction} of the form $(i, \lambda, \side; i', \lambda', \side')$ where
\myitems{
    \item $i, i' \in [1, k]$;
    \item $\lambda \in [n_i], \lambda' \in [n_{i'}]$;
    \item $\side$ and $\side'$ are functions chosen from the set $\{\xleft, \xright, \ybot$, $\ytop\}$.
}
Each leaf is associated with an {\em result}, which is defined to be a subset of $[n_1] \times [n_2] \times ... \times [n_k]$.

@@

Given an $(n_1, n_2, ..., n_k)$-input $\set{R_1, R_2, ..., R_k}$, we execute the algorithm by traversing a root-to-leaf path in $\T$. In general, at an internal node with the instruction $(i, \lambda, \side; i', \lambda', \side')$, we first identify
\myitems{
    \item $r$ as the rectangle with ID $\lambda$ in $R_i$;
    \item $r'$ as the rectangle with ID $\lambda'$ in $R_i'$.
}
Then, we compare $\side(r)$ to $\side'(r')$ (e.g., if $\side = \xright$ and $\side = \ybot$, then the comparison is between $\xright(r)$ to $\ybot(r')$). We branch to
\myitems{
    \item the left child if $\side(r) < \side'(r')$;
    \item the middle child if $\side(r) = \side'(r')$;
    \item the right child if $\side(r) > \side'(r')$.
}
At a leaf, we look at its associated result and, for each $(\lambda_1, \lambda_2, ..., \lambda_k)$ in the result, output $(r_1, r_2, ..., r_k)$ where $r_i$ ($i \in [k]$) is the rectangle with ID $\lambda_i$ in $R_i$.

An algorithm is {\em correct} if it produces the correct output on all $(n_1, n_2, ..., n_k)$-inputs. Its {\em worst-case cost} is the height of $\T$.
